Canopy leaf area of an evergreen Eucalyptus woodland does not respond to elevated

### 2015-05-01 manuscript.Rmd
references

The soil at the site is a loamy sand in the Clarendon Formation (Barton et al., 2010; Isbell, 2002), a slightly-acid, low-organic C, well-drained soil. The soil surface is a loamy sand with > 75% sand content, and from ca. 50 to 300 cm depth, soils are a sandy clay loam with > 30% silt and clay. The soil is nutrient poor, and a strong limitation of P on tree growth at the site has recently been demonstrated (Crous et al., 2015). Groundwater depth is 12.6 – 12.9 m below the soil surface. The woodland ranges in density from 600-1000 trees ha-1. The vegetation consists of an open woodland (basal area before treatment, BA = `r mean(ba$BA)`, SE = `r sd(ba$BA)/sqrt(6)` m$^2$ ha$^{-1}$, n = 6 plots) with a single major canopy-forming tree (*Eucalyptus tereticornis* Sm.) with a minor component of *E. amplifolia* Naudin. The understory stratum is composed mainly of grass species dominated by Microlaena stipoides (Labill.) R.Br., with lower densities of forbs of *Commelina cyanea* R.Br., *Lobelia purpurascens* R.Br., *Dichondra repens* J.R.Forst. & G.Forst and occasional shrubs of *Breynia oblongifolia* (Mull.Arg.) Mull.Arg. and *Bursaria spinosa* Cav.

Within the EucFACE site, there are six 25-m diameter circular plots (designated hereafter as ‘rings’). Each ring comprises 32 vent pipes in a cylindrical frame extending above the tree canopy to 28 m high. Each vertical pipe is subtended by a control valve and is connected to a circular plenum (Hendrey et al., 1999) through which a computer-controlled system modulates release of diluted $CO_2$ into the vegetation according to a proportional-integral-differential type of algorithm (Lewin et al., 2009). The vegetation within three randomly selected rings was exposed to a $CO_2$ mole fraction ($C_a$) higher than ambient starting in September 2012, whereas the other three rings were identical to the $CO_2$ treatment rings in infrastructure and instrumentation but received ambient $C_a$, and are used as control plots. The target diurnal $C_a$ for the rings exposed to $eC_a$ was 150 μmol $CO_2$ mol$^{-1}$ above ambient $C_a$ in full operation mode. However, $C_a$ was increased gradually for six months at a rate of ca. 30 μmol $CO_2$ mol$^{-1}$ month$^{-1}$ until full operation mode was reached (Figure 1a).

@@

## Leaf litter production
In each ring, 8 litter cloth baskets with an area of 0.1979 m$^2$ were installed at 0.5m above the forest floor. Litter was collected approximately monthly, and leaves separated from other components such as twigs, bark, seeds, insect frass, etc. Litter production was converted to units of *L* using an estimate of specific leaf area (SLA) determined from 329 leaves collected from the tree canopies from May 2012 to May 2014 (`r get_sla()` cm$^2$ g$^{-1}$, SE=2.14). We did not find a significant effect of $C_a$ treatment on SLA (Ellsworth, unpublished data).

## Estimation of leaf area index from digital photography
Leaf area index (*L*) was estimated from zenith canopy photographs following (Macfarlane et al., 2007). From October 2012 to October 2013, we took ca. monthly canopy photographs in each FACE ring. Photographs were taken in the morning on cloudy days. We took one photograph at each point of a 4 x 4 m grid established within each ring, unless obstructed by trees, yielding 20-23 photos per ring. Sampling points too close to either the vertical vent pipes (edges of the ring) or the central scaffolding tower were excluded. The camera was aligned with the north-south axis and set at 1.25 m height.  Camera lens was pointed directly upwards and levelled using a three axis bubble level. We used a Nikon D90 camera with a Nikon AF 50 mm lens (Nikon Corporation, Tokyo, Japan). To measure cover, the camera was set to automatic exposure, Aperture-Priority and automatic autofocus mode and F10 lens. Following recommendations by Jonckheere et al. (2005), we used the automated Ridler algorithm to threshold the canopy photos. We then calculated canopy transmittance (τ_p) from each canopy photo as the proportion of pixels with intensity in the blue band below that threshold. *L* from canopy photos (Lp) was calculated with a simple Beer-Lambert law (Eq. 1).

>>>>

The soil at the site is a loamy sand in the Clarendon Formation (Barton et al., 2010; Isbell, 2002), a slightly-acid, low-organic C, well-drained soil. The soil surface is a loamy sand with > 75% sand content, and from ca. 50 to 300 cm depth, soils are a sandy clay loam with > 30% silt and clay. The soil is nutrient poor, and a strong limitation of P on tree growth at the site has recently been demonstrated [@Crous2015]. Groundwater depth is 12.6 – 12.9 m below the soil surface. The woodland ranges in density from 600-1000 trees $ha^{-1}$. The vegetation consists of an open woodland (basal area before treatment, BA = `r mean(ba$BA)`, SE = `r sd(ba$BA)/sqrt(6)` m$^2$ ha$^{-1}$, n = 6 plots) with a single major canopy-forming tree (*Eucalyptus tereticornis* Sm.) with a minor component of *E. amplifolia* Naudin. The understory stratum is composed mainly of grass species dominated by *Microlaena stipoides* (Labill.) R.Br., with lower densities of forbs of *Commelina cyanea* R.Br., *Lobelia purpurascens* R.Br., *Dichondra repens* J.R.Forst. & G.Forst and occasional shrubs of *Breynia oblongifolia* (Mull.Arg.) Mull.Arg. and *Bursaria spinosa* Cav.

Within the EucFACE site, there are six 25-m diameter circular plots (designated hereafter as ‘rings’). Each ring comprises 32 vent pipes in a cylindrical frame extending above the tree canopy to 28 m high. Each vertical pipe is subtended by a control valve and is connected to a circular plenum [@Hendrey1999] through which a computer-controlled system modulates release of diluted $CO_2$ into the vegetation according to a proportional-integral-differential type of algorithm [@Lewin2009]. The vegetation within three randomly selected rings was exposed to a $CO_2$ mole fraction ($C_a$) higher than ambient starting in September 2012, whereas the other three rings were identical to the $CO_2$ treatment rings in infrastructure and instrumentation but received ambient $C_a$, and are used as control plots. The target diurnal $C_a$ for the rings exposed to $eC_a$ was 150 μmol $CO_2$ mol$^{-1}$ above ambient $C_a$ in full operation mode. However, $C_a$ was increased gradually for six months at a rate of ca. 30 μmol $CO_2$ mol$^{-1}$ month$^{-1}$ until full operation mode was reached (Figure 1a).

@@

## Leaf litter production
In each ring, 8 litter cloth baskets with an area of 0.1979 $m^2$ were installed at 0.5m above the forest floor. Litter was collected approximately monthly, and leaves separated from other components such as twigs, bark, seeds, insect frass, etc. Litter production was converted to units of *L* using an estimate of specific leaf area (SLA) determined from 329 leaves collected from the tree canopies from May 2012 to May 2014 (`r get_sla()` cm$^2$ g$^{-1}$, SE=2.14). We did not find a significant effect of $C_a$ treatment on SLA (Ellsworth, unpublished data; Ben Moore, unpublished data).

## Estimation of leaf area index from digital photography
Leaf area index (*L*) was estimated from zenith canopy photographs following [@Macfarlane2007]. From October 2012 to October 2013, we took ca. monthly canopy photographs in each FACE ring. Photographs were taken in the morning on cloudy days. We took one photograph at each point of a 4 x 4 m grid established within each ring, unless obstructed by trees, yielding 20-23 photos per ring. Sampling points too close to either the vertical vent pipes (edges of the ring) or the central scaffolding tower were excluded. The camera was aligned with the north-south axis and set at 1.25 m height.  Camera lens was pointed directly upwards and levelled using a three axis bubble level. We used a Nikon D90 camera with a Nikon AF 50 mm lens (Nikon Corporation, Tokyo, Japan). To measure cover, the camera was set to automatic exposure, Aperture-Priority and automatic autofocus mode and F10 lens. Following recommendations by Jonckheere et al. [-@Jonckheere2005], we used the automated Ridler algorithm to threshold the canopy photos. We then calculated canopy transmittance (τ_p) from each canopy photo as the proportion of pixels with intensity in the blue band below that threshold. *L* from canopy photos (Lp) was calculated with a simple Beer-Lambert law (Eq. 1).
